---
title: "LAS 6292: PAPERLESS DATA COLLECTION"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * Reminder to install epicollect app on phone or tablet

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* Be able to explain the advantages of paperless data collection
* Identify means of preventing data loss when collecting data electronically
* Learn how to build a form, and collect, upload, and download data using EpiCollect
  
## Pre-class Preparation (Students):

Download the Epicollect app to a phone or tablet (if you need one let me know).

* [Android](https://play.google.com/store/apps/details?id=uk.ac.imperial.epicollect.five&hl=en_US&gl=US)

* [Apple](https://apps.apple.com/us/app/epicollect5/id1183858199)

### **Readings**

1.  Aanensen DM, Huntley DM, Feil EJ, al-Own F, Spratt BG (2009) EpiCollect: Linking Smartphones to Web Applications for Epidemiology, Ecology and Community Data Collection. PLoS ONE 4(9): e6968.  
        [[read online]](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0006968) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/instructor-materials/Aanensen_etal_2009.PDF)

2.  Moylan, CA et al. 2013. Increasingly mobile: How new technologies can enhance qualitative research. Qualitative social work: research and practice, 14(1):36-47.  
        [[read online]](https://journals.sagepub.com/doi/10.1177/1473325013516988) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/instructor-materials/Moylan_etal_2015.pdf)

3.  Teacher, Amber G. F. et al. Smartphones in ecology and evolution: a guide for the app‐rehensive. Ecology and Evolution 3(16):5268– 5278  
        [[read online]](https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.888)  [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/instructor-materials/Teacher_etal_2013.pdf)
        

### **Online Lectures:**  

* [Paperless Data Collection](https://youtu.be/n8aoe3pXCGU)  

* Great tutorials by UConn's David Dickson  

    * [Intro to EpiCollect 5](https://kaltura.uconn.edu/playlist/dedicated/1_14lpwsw0/1_fmj06huc)  
            
    * [Creating an EpiCollect PROJECT](https://kaltura.uconn.edu/media/Creating+an+Epicollect+5+Project/1_853ybdbf)  
        
    * [Creating a FORM for your Epicollect Project](https://kaltura.uconn.edu/media/Creating+a+Form+in+Epicollect+5/1_inr7wxq8)  
        
    * [Collecting data in EpiCollect with your phone or tablet](https://kaltura.uconn.edu/media/Collecting+Data+With+Epicollect+5/1_kgzfddtp)  
        
        

## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



### **Topic 1 Overview: Paperless Data Collection** (`r t1_overview` min)

### Breakout

**Breakout (`r breakout1` min): ** Paperless data collection with Epicollect


### Break (`r snack_break` min)

### Return and discuss breakout results

1. **Returning results & Take-home message (`r breakout1_return` min)** 

2. **Take-home message:** Reinforce value of paperless data collection
      
### Free Time

There are `r time_remain` min remaining that can be used to -----

# Tools & Resources

1. [EpiCollect user's guide](https://docs.epicollect.net/)

2. Best data app for interactive voice data collection: [Magpi](https://www.magpi.com/blog/interactive-data-collection) (Web, iOS, Android)

3. Lists of useful apps for data collection in the field: 

    * [list 1](https://www.teamscopeapp.com/mobile-data-collection-guide/7-mobile-data-collection-apps-for-field-research)

    * [list 2](http://brunalab.org/apps/ )

    * [list 3](https://www.researchgate.net/post/Please-could-you-indicate-some-useful-android-apps-for-Ecology-Forestry-or-Environment)

    * [list 4](https://writingfornature.wordpress.com/2012/08/17/tools-for-the-modern-primate-apps-for-field-work/)

    * [list 5](https://ayrnstories.wordpress.com/2017/08/09/apps-in-ecological-research/)

    * [list 6](https://www.commonsense.org/education/top-picks/excellent-ecology-and-environmental-science-apps-games-and-websites)

    * [list 7](https://www.researchgate.net/post/What_essential_android_apps_should_a_field_biologist_know_about)

# Sources

1. [Keep calm and go paperless: Electronic lab notebooks can improve your research](https://openworking.wordpress.com/2019/07/05/keep-calm-and-go-paperless-electronic-lab-notebooks-can-improve-your-research/)

2. [NOAA Pathways to paperless Data Collection](https://www.fisheries.noaa.gov/feature-story/pathway-paperless-data-collection)

3. [Thieler ER, Zeigler SL, Winslow LA, Hines MK, Read JS, Walker JI (2016) Smartphone-Based Distributed Data Collection Enables Rapid Assessment of Shorebird Habitat Suitability. PLoS ONE 11(11): e0164979.](https://doi.org/10.1371/journal.pone.0164979)

4. [Teacher, A. G., Griffiths, D. J., Hodgson, D. J., & Inger, R. (2013). Smartphones in ecology and evolution: a guide for the app-rehensive. Ecology and evolution, 3(16), 5268–5278.](https://doi.org/10.1002/ece3.888)

5. [Palumbo, M.J., Johnson, S.A., Mundim, F.M., Lau, A., Wolf, A.C., Arunachalam, S., Gonzalez, O., Ulrich, J.L., Washuta, A. and Bruna, E.M. (2012), Harnessing Smartphones for Ecological Education, Research, and Outreach. The Bulletin of the Ecological Society of America, 93: 390-393.](https://esajournals.onlinelibrary.wiley.com/doi/10.1890/0012-9623-93.4.390)



# DECIDE WHAT TO INCLUDE ON RESOURCES FOR SESSION AND COMMENT OUT THE REST




## **Week 11: ‘Paperless’ data collection**  

https://www.linkedin.com/learning/google-sheets-advanced-tips-and-tricks/use-ifttt-with-google-sheets?u=41282748


*  sometimes hacking is worth it to avoid writing. [calipers-to-excel](https://www.notion.so/Hacking-Digital-Calipers-3ee7726f11ca431694dc70a1977516e4)  

*  Mendi Benigni. Using Google Forms, Google Sheets & Excel to Collect and Analyze Data: http://tinyurl.com/ycnm8svv
*  https://medium.com/google-design/the-evolution-of-material-designs-text-fields-603688b3fe03 
Readings
*  https://www.smashingmagazine.com/2018/08/best-practices-for-mobile-form-design/
*  http://www.kwantu.net/blog/2016/12/26/best-practices-for-designing-data-collection-forms   
*  barcode software (cheap or free) https://www.barcodesinc.com/cats/barcode-scanners/apps.htm
*  Rayhan, R. U., Zheng, Y., Uddin, E., Timbol, C., Adewuyi, O., & Baraniuk, J. N. (2013). Administer and collect medical questionnaires with Google documents: a simple, safe, and free system. Applied Medical Informatics, 33(3), 12-21.  
*  Moylan, C. A., Derr, A. S., & Lindhorst, T. (2013). Increasingly mobile: How new technologies can enhance qualitative research. Qualitative social work: research and practice, 14(1), 36-47.  
*  Aanensen DM, Huntley DM, Feil EJ, al-Own F, Spratt BG (2009) EpiCollect: Linking Smartphones to Web Applications for Epidemiology, Ecology and Community Data Collection. PLoS ONE 4(9): e6968.  
*  Harari, G. M., Lane, N. D., Wang, R., Crosier, B. S., Campbell, A. T., & Gosling, S. D. (2016). Using Smartphones to Collect Behavioral Data in Psychological Science: Opportunities, Practical Considerations, and Challenges. Perspectives on Psychological Science 11(6), 838-854.  
*  tropy for photos looks fantastic, can export data as json, csv etc https://tropy.org/  




