---
title: "LAS 6292: AUTOMATED DATA EXTRACTION"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* OCR 
* 
  
## Pre-class Preparation (Students):

### **Readings**


### **Online Lectures:**  

  * [NIH:  Statistical methods for protecting data confidentiality](https://www.youtube.com/watch?v=3vN-6A2zbsk)
Patricia Jenkinson Video Seminar on [Ethics: Human Subject Research](https://www.youtube.com/watch?v=e0uydNe4cQo)
  * [An Overview of NIH Policies on Human Subjects Research](https://www.youtube.com/watch?v=jPWxPgR-vww) Leisten at 1.75 speed
  * [U.S. Department of Health and Human Services: Informed COnsent Overview](https://www.youtube.com/watch?v=Y7uI3sM9wtc)
  * Australian Council for International Development (ACFID): [Principles and Guidelines for Ethical Research and Evaluation in Development](https://www.youtube.com/watch?v=ylkKCgEIMws)

  
## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



### **Topic 1 Overview: Annonymization. Also MEMES!!** (`r t1_overview` min)
https://amnesia.openaire.eu/index.html
https://www.nist.gov/itl/applied-cybersecurity/privacy-engineering/collaboration-space/focus-areas/de-id/tools
https://www.techfunnel.com/information-technology/3-types-of-data-anonymization-techniques-and-tools-to-consider/ 
https://www.educause.edu/focus-areas-and-initiatives/policy-and-security/cybersecurity-program/resources/information-security-guide/toolkits/guidelines-for-data-deidentification-or-anonymization

https://ori.hhs.gov/education/products/columbia_wbt/rcr_data/case/index.html#2

https://github.com/BrunaLab/BDFFP-precipitation/issues/27 

Guidelines: it had to be original, had to illustrate something they learned 
in class, and had to use the meme format more or less properly.

https://twitter.com/v_matzek/status/1376959852367372288 
https://www.iloveimg.com/meme-generator 
https://www.kapwing.com/meme-maker
https://memegenerator.net/create 

### *Topic 2 Overview: Topic*  (`r t2_overview` min)

Intro text

### Breakout & Return Results 

**Breakout (`r breakout1` min): ** topic of breakout 

**Returning results & Take-home message (`r breakout1_return` min)** 
summary of results

2. **Take-home message:** message.

additional text
      
### anything before the break? (`r cleaning_demo` min)

if so, describe here    
    
### Break (`r snack_break` min)

### Anything after the break? (`r proj_overview` min)

If so describe here


### Breakout 2: In-class Assignment (`r breakout2` min)
Text about the breakout


### Free Time

There are `r time_remain` min remaining that can be used to -----


# Tools & Resources

Anonymizing Tools
i.	https://windowsreport.com/data-anonymization-software/ 
ii.	https://aircloak.com/top-5-free-data-anonymization-tools/ 
