---
title: "LAS 6292: LEGAL AND ETHICAL ISSUES"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* Be able to explain the major laws, policies, and ethical guidelines relevant to data collection & management
* Become familiar with relevant concepts and terminology (copyrights, licenses, and waivers)
* Explain the difference between Restricted, Sensitive, and open Data
* Discuss ethical considerations surrounding the use, reuse, and sharing of research data
* Describe the tools and processes available for anonymizing data. 
  
## Pre-class Preparation (Students):

### **Readings**

1.  de Koning M, Meyer B, Moors A, Pels P. 2019. Guidelines for anthropological research: Data management, ethics, and integrity. Ethnography. 20(2):170-174.  [read online](https://journals.sagepub.com/doi/10.1177/1466138119843312)  [download pdf](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Koning_etal_2019.pdf)

2.  Katie Fortney, UC Office of Scholarly Communication. 2016. “Who owns your data?”  [read online](https://uc3.cdlib.org/2016/09/08/who-owns-your-data/) [download pdf](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Fortney_2016.pdf)

3. Goodluck, K. 2020. Indigenous data sovereignty shakes up research. High Country News.  [read online](https://www.hcn.org/issues/52.11/indigenous-affairs-covid19-indigenous-data-sovereignty-shakes-up-research) 

4. Woodbury, R. B., Beans, J. A., Hiratsuka, V. Y., & Burke, W. (2019). Data Management in Health-Related Research Involving Indigenous Communities in the United States and Canada: A Scoping Review. Frontiers in genetics, 10, 942. [read online](https://www.hcn.org/issues/52.11/indigenous-affairs-covid19-indigenous-data-sovereignty-shakes-up-research)]  [download pdf](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Goodluck_2020.pdf)

### **Online Lectures:**  

  * Quick Overview [Video Lecture by EB](https://youtu.be/sBWzs3Ra5fw)
  * [NIH:  Statistical methods for protecting data confidentiality](https://www.youtube.com/watch?v=3vN-6A2zbsk)
  * Patricia Jenkinson Video Seminar on [Ethics: Human Subject Research](https://www.youtube.com/watch?v=e0uydNe4cQo)
   * [U.S. Department of Health and Human Services: Informed COnsent Overview](https://www.youtube.com/watch?v=Y7uI3sM9wtc)
  * Australian Council for International Development (ACFID): [Principles and Guidelines for Ethical Research and Evaluation in Development](https://www.youtube.com/watch?v=ylkKCgEIMws)
 * [An Overview of NIH Policies on Human Subjects Research](https://www.youtube.com/watch?v=jPWxPgR-vww) Optional, listen at 1.75 speed!
  
## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```

### Before we start: 

1. [example](https://github.com/BrunaLab/BDFFP-precipitation/issues/27) of a real data clean-up situation for which OpenRefine is ideal


### **Topic 1 Overview: Anonymization.** (`r t1_overview` min)

1. [Guidelines for Data De-Identification or Anonymization](https://www.educause.edu/focus-areas-and-initiatives/policy-and-security/cybersecurity-program/resources/information-security-guide/toolkits/guidelines-for-data-deidentification-or-anonymization)

### Breakout & Return Results 

**Breakout (`r breakout1` min): ** discussion of case studies

Case study for breakout: [who owns the data?](https://ori.hhs.gov/education/products/columbia_wbt/rcr_data/case/index.html#2)


**Returning results & Take-home message (`r breakout1_return` min)** 

1. summary of discussion

2. **Take-home message:** message.

      
### anything before the break? (`r cleaning_demo` min)

### Break (`r snack_break` min)

### *Topic 2 Overview: MEME-O-RAMA*  (`r t2_overview` min)

*Guidelines: *has to be original, has to illustrate something learned 
in class, and have to use the meme format more or less properly. Foreign Lang Memes encouraged.

1. examples from a stats class: [link](https://twitter.com/v_matzek/status/1376959852367372288 )
2. Meme generators: 
    * [meme generator 1](https://www.iloveimg.com/meme-generator)
    * [meme generator 2](https://www.kapwing.com/meme-maker) 
    * [meme generator 3](https://memegenerator.net/create)



### Free Time

There are `r time_remain` min remaining that can be used to -----


# Tools & Resources

1. Anonymizing Tools
    * [list 1](https://windowsreport.com/data-anonymization-software/)
    * [list 2](https://aircloak.com/top-5-free-data-anonymization-tools/) 
    * [Amnesia (data anonymization algorithms)](https://amnesia.openaire.eu/index.html)
    * De-identification Tools from [NIST](https://www.nist.gov/itl/applied-cybersecurity/privacy-engineering/collaboration-space/focus-areas/de-id/tools)
    * [3 Types of Data Anonymization Techniques and Tools to Consider](https://www.techfunnel.com/information-technology/3-types-of-data-anonymization-techniques-and-tools-to-consider/)
 
 2. Guidelines
    * [Guidelines for Data De-Identification or Anonymization](https://www.educause.edu/focus-areas-and-initiatives/policy-and-security/cybersecurity-program/resources/information-security-guide/toolkits/guidelines-for-data-deidentification-or-anonymization)
