---
title: "LAS 6292: Data Organization in Spreadsheets"
date: "updated: `r Sys.Date()`"
header-includes:
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
- \usepackage[default]{sourcesanspro}
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
# month & year on header left
- \fancyhead[L]{LAS 6292 Spring `r format(Sys.time(), '%Y')`, EM Bruna} 
# name and page on header right
- \fancyhead[R]{QA/QC 2 - OpenRefine, p. \thepage} 
- \fancyfoot[R]{Last update `r format(Sys.time(), '%d %B %Y')`}
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
- \usepackage{xcolor} # changes colors of text
# to center to top level (#) headings (my name) and change color of font
# more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# to center to sub-sub-headings (###) and change color of font
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# - \usepackage{ragged2e}
# - \setlength\RaggedRightParindent{15pt}
# - \RaggedRight - # make the right margin ragged
# https://tex.stackexchange.com/questions/258714/prevent-hyphenation-across-lines-enforce-right-margin
# - \hyphenpenalty=10000 #to suppress ordinary hyphenation
# - \exhyphenpenalty=10000 #to prohibit line breaks at explicit hyphens.
# - \setstretch{1.0}

# https://babichmorrowc.github.io/post/changing-fonts-in-rmarkdown-pdfs/

mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  # html_document:
  #   df_print: paged
  # rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Class Outline: QA/QC 2 - Open Refine

## Objectives and Competencies:   

By the end of this lesson students will:

## Pre-Class Preparation (Instructor):  

**Bring to Class:  **  

  * Snacks
  * Tent cards for student names

## Pre-class Preparation (Students):

### **Online Lectures:**  None

### **Readings**: 

1. [“Ch7 in Exploratory Data Analysis” in Wickham and Grolemund’s “R for Data Science”:](https://r4ds.had.co.nz/exploratory-data-analysis.html)  

2. Rougier NP, Droettboom M, Bourne PE (2014) Ten Simple Rules for Better Figures. PLoS Comput Biol 10(9): e1003833. [read online](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003833)

### **Online Lectures:**  

1. Tommy McCall: [The simple genius of a good graphic](https://www.ted.com/talks/tommy_mccall_the_simple_genius_of_a_good_graphic?language=en)
    
2. R. Luke DuBois: [Insightful human portraits made from data](https://www.ted.com/talks/r_luke_dubois_insightful_human_portraits_made_from_data?language=en)

3. Mike Chapple: [Data Visualization in R with ggplot2](https://www.linkedin.com/learning/data-visualization-in-r-with-ggplot2/the-grammar-of-graphics?resume=false&u=41282748). Watch the "Introducing ggplot2" (4:43) and "Grammar of Graphics" (4:54)




### Computer Resources


```{r timer, echo=FALSE}
part_1 <- 10
part_2 <- 10
part_3 <- 45
snack_break <- 10
part_4 <-30
part_5 <-45
part_6 <-10

total_time <- 
  part_1 + 
  part_2 + 
  part_3 + 
  part_4 +
  part_5 +
  part_6 +
  snack_break 

time_remain <- 180 - total_time
```

## In-Class: Live cofing demo of summary and graphics


1. Intro slides

2. Load CSV file of df

3. Use code here for demo: [LAS DEMO CODE](https://gist.github.com/embruna/d0a42a5eac5fd56b582fe29d0fd26259)

### **Intro to OR** (`r part_1` min)

### **Working with OR**  (`r part_2` min)  

### **Filtering and Sorting** (`r part_3` min)
     
### Break (`r snack_break` min)

### **Examining Numbers** (`r part_4` min)

### **Using Scripts, Exporting, and Saving**  (`r part_5` min)  

### **Wrap-up, Questions**  (`r part_6` min)  

## **Assignment** (`r time_remain` min)


6. **Submission:** Submit a pdf of your image as `week7_hw` on Canvas.  

### Grading Rubric: 

Data corrected and JSON file can be used on another data set: 50  
Most data correction properly programmed; some require instructor follow-up: 40  
Many of the corrections missing, JSON file unable to process new data : 30  
Instructor follow-up required to implement most changes: 20  

# Sources for this lesson 


# Additional Tools and Resources

## Tutorials


## R Tools

https://rich-iannone.github.io/pointblank/

package validate

https://www.yan-holtz.com/PDF/Dataviz_Caveats.pdf


Weissgerber TL, Milic NM, Winham SJ, Garovic VD (2015) Beyond Bar and Line Graphs: Time for a New Data Presentation Paradigm. PLoS Biol 13(4): e1002128. https://doi.org/10.1371/journal.pbio.1002128
