---
title: "LAS 6292: TRANSCIRPTION & TRANSLATION"
date: "updated: `r Sys.Date()`"
header-includes:
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
- \usepackage[default]{sourcesanspro}
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
# month & year on header left
- \fancyhead[L]{LAS 6292 Spring `r format(Sys.time(), '%Y')`, EM Bruna} 
# name and page on header right
- \fancyhead[R]{Transcription and Translation, p. \thepage} 
- \fancyfoot[R]{Last update `r format(Sys.time(), '%d %B %Y')`}
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
- \usepackage{xcolor} # changes colors of text
# to center to top level (#) headings (my name) and change color of font
# more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# to center to sub-sub-headings (###) and change color of font
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# - \usepackage{ragged2e}
# - \setlength\RaggedRightParindent{15pt}
# - \RaggedRight - # make the right margin ragged
# https://tex.stackexchange.com/questions/258714/prevent-hyphenation-across-lines-enforce-right-margin
# - \hyphenpenalty=10000 #to suppress ordinary hyphenation
# - \exhyphenpenalty=10000 #to prohibit line breaks at explicit hyphens.
# - \setstretch{1.0}

# https://babichmorrowc.github.io/post/changing-fonts-in-rmarkdown-pdfs/

mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  # html_document:
  #   df_print: paged
  # rmdformats::robobook: default
---



```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
knitr::opts_chunk$set(echo = TRUE)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```



## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * Ask for volunteers to record text in new languages.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

  * Understand the different approaches to translation and the research questions for which each is most appropriate
  * Become familiar with software for automated translation via direct upload and api
  * Learn to use the different types of automated and manual transcription tools
  * Understand the strengths and weaknesses of automated translation 
  * Be able to describe the metadata required for transcription files

## Pre-class Preparation (Students):

### **Readings**


1.  Bakker, Rebecca. "Transcription Tools and Software" (2017). Works of the FIU Libraries. 62. [[read online]](https://digitalcommons.fiu.edu/glworks/62) [download pdf](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Bakker_transcription_tools_slides.pdf)  

2. McMullin, C. 2021, Transcription and Qualitative Methods: Implications for Third Sector Research. _Voluntas._ [link](https://doi.org/10.1007/s11266-021-00400-3)


    _OPTIONAL (but very, very useful)_

1. Daniel G. Oliver, Julianne M. Serovich, Tina L. Mason. 2005. Constraints and Opportunities with Interview Transcription: Towards Reflection in Qualitative Research, _Social Forces_ 84(2): 1273–1289 [link](https://doi.org/10.1353/sof.2006.0023)

1. Bailey, J. 2008. First steps in qualitative data analysis: transcribing. _Family Practice_ 25(2):127–131 [link](https://doi.org/10.1093/fampra/cmn003)

### **Online Lectures:**  

1. Dr. Jarek Kriukow: [How to transcribe interviews - Part 1: "naturalism" and "denaturalism"](https://www.youtube.com/watch?v=NzVjlKR4HQQ) and [How to transcribe interviews - Part 2: Which approach to use?](https://www.youtube.com/watch?v=69T8aVf5Nj0)

    _Note: we will discuss anonymization during the session on legal and ethical Issues_

## Class Outline

```{r timer, echo=FALSE}
t1_overview <- 20
breakout1 <- 15
breakout1_return <-15
breakout2 <- 15
breakout2_return <-15
t2_overview <- 30
snack_break <- 10
breakout2 <- 45

total_time <- 
  t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  breakout2 + 
  breakout2_return + 
  snack_break + 
  breakout2 

time_remain <- 180 - total_time
```


### **Overview: Transcription / Translation** (`r t1_overview` min)

### **Breakout 1: Translation** (`r breakout1` min)

### **Returning results 1** (`r breakout1_return` min)

### Break (`r snack_break` min)

### **Breakout 2: Transcription** (`r breakout2` min)

### **Returning results 2** (`r breakout2_return` min)


### Free Time

There are `r time_remain` min remaining that can be used to -----

# Tools & Resources for Transcription & Translation

## Translation

### Text: Automated/AI
        
1. [Google Translate](https://translate.google.com/) and [help documents](https://support.google.com/docs/answer/4492226?hl=en#zippy=%2Cadd-punctuation).

1. Google Docs: [in-doc translation](https://support.google.com/docs/answer/187189?co=GENIE.Platform%3DDesktop&hl=en). See also how to [translate using google sheets](https://twitter.com/CienciaDelCope/status/1203384707708325889)
 
1. MS Word: [in-doc translation](https://support.microsoft.com/en-us/topic/translate-text-into-a-different-language-287380e4-a56c-48a1-9977-f2dca89ce93f) 

1. [DeepL](https://www.deepl.com/translator)  

### Audio: Automated/AI  

1. [Happy Scribe](https://www.happyscribe.com/). 40+ languages. 30 min. free then paid; [student pricing available](https://www.happyscribe.com/audio-to-text-for-students). Can integrate with Zapier to up/download files from [Google Drive](https://zapier.com/apps/happy-scribe/integrations/google-drive) and [Dropbox](https://zapier.com/apps/happy-scribe/integrations/dropbox) **_Multiple languages_**
  
  
1. Google Audio Translation [in Beta](https://cloud.google.com/translate/media/docs)
### Text or Audio: API or Command Line

1. [```googleLanguageR```](https://docs.ropensci.org/googleLanguageR/). Uses the Google Cloud machine learning APIs for text and speech tasks. The [Cloud Translation API](https://cloud.google.com/translate/) does detection and translation of text. 

1.  [```aws.translate```](https://github.com/cloudyr/aws.translate) uses the Amazon Web Services to do transcription via API. Has program for NGO/non-profits.

1. [```translateR``` package](https://cran.r-project.org/web/packages/translateR/index.html): tools to access Google Cloud Translation via API. [Overview](https://www.google.com/search?q=translation+in+R&rlz=1C5CHFA_enUS828US828&oq=translation+in+R&aqs=chrome..69i57j0i20i263i433i512j0i131i433i512j0i20i263i512j0i512j0i433i512j0i131i433i512j0i512l3.2043j0j7&sourceid=chrome&ie=UTF-8) and [reference manual](https://cran.r-project.org/web/packages/translateR/translateR.pdf)

1. [Watson (IBM)](https://cloud.ibm.com/catalog/services/speech-to-text) 500 minutes free

1. [DeepL](https://www.deepl.com/translator)

1. [Translate Shell](https://www.soimort.org/translate-shell/#translate-shell) via API

1. [Translate Shell](https://www.soimort.org/translate-shell/#translate-shell)

### Text: Manual with Computer Assisted Technology (CAT) Software

1. Video overview of different  [CAT Tools](https://www.youtube.com/watch?v=nbxBV_7QEyk) from 'Freelanceverse'.

1. [OmegaT CAT Software](https://omegat.org/). (Free)

1. [WordFast Anywhere](https://www.wordfast.com/products_wordfast_anywhere) (free online; desktop version are paid)

1. [Across.net](https://www.across.net/) (basic version free)

## Transcription

1.  [Comparison Video](https://www.youtube.com/watch?v=GAmkIHJl7A0 comparisons)

2. NY Times' Wirecutter [review of transcription services](https://www.nytimes.com/wirecutter/reviews/best-transcription-services/) 

### Automated via upload

  **_Top-Tier _**  

1. [Otter.ai](https://otter.ai/login). English only; free (600 min/mo) and paid plans. Can transcribe zoom meetings (e.g., if you do interviews via zoom) and can [integrate with Dropbox](https://blog.otter.ai/otter-dropbox-integration/). Phone app is available for [iOS](https://play.google.com/store/apps/details?id=com.aisense.otter&hl=en_US&gl=US). There is an [online tutorial](https://www.youtube.com/watch?v=SCaBnObVsqI).  **_English-only_**
  
1.  [Happy Scribe](https://www.happyscribe.com/) - see descriptionabove.

1.  [Temi](https://www.temi.com/). **_English Only_**, less accurate than Otter.ai and Happy Scribe.

1.  Online (free demo) version of [Watson](https://speech-to-text-demo.ng.bluemix.net/); note the "detect multiple speakers" option.

1. [Zoom via UF](https://video.ufl.edu/conferencing/zoom/captions-and-live-transcription/); see also this [Link](https://education.ufl.edu/educational-research/14320).

  **_Second-Tier_**  

1.  [Spext](https://www.spext.co/). Speech-audio-text editing; free and paid plans.

1. [Amazon Transcribe](https://aws.amazon.com/transcribe/)

### Automated via API

1. [```transcribeR``` package](https://cran.r-project.org/web/packages/transcribeR/index.html) uses HP IDOL OnDemand's Speech Recognition API.                * [Vignette](https://cran.r-project.org/web/packages/transcribeR/vignettes/Transcribing_audio_with_transcribeR.html) 
    * [Reference manaual](https://cran.r-project.org/web/packages/transcribeR/transcribeR.pdf)
    * [tutorial](http://www.micha-dietze.de/pages/r_texts/_book/kaleidoscope.html)

1. [```googleLanguageR```](https://docs.ropensci.org/googleLanguageR/). Uses the Google Cloud machine learning APIs for text and speech tasks. The [Cloud Speech API](https://cloud.google.com/speech/) transcribes sound files to text and the Cloud Text-to-Speech AP

1. [```AWS transcribe```](https://github.com/cloudyr/aws.transcribe) uses the Amazon Web Services to do transcription via API. Has program for NGO/non-profits.

### Automated via dictation

1. [Google Docs](https://www.google.com/docs). See also [Google Docs Voice Recognition](https://support.google.com/docs/answer/4492226?hl=en) and [helpful how-to blog post on transcription using google docs and voice recognition](https://qz.com/work/1087765/how-to-transcribe-audio-fast-and-for-free-using-google-docs-voice-typing/)  

1. [SpeechNotes](https://speechnotes.co/)

1. [Apple Dictation](https://support.apple.com/en-us/HT202584) and [Dictation Commands](https://support.apple.com/en-gb/HT203085)

1. [How to use Windoes Voice Dictation](https://www.windowscentral.com/how-use-voice-dictation-windows-10) and [Windows Speech Recognition Commands](https://support.microsoft.com/en-us/help/12427/windows-speech-recognition-commands)

### Manual with Software 

1. [oTranscribe](https://otranscribe.com/)  Incredibly easy, web based, with time stamps, variable speed.

1. [Express Scribe](Transcription Software https://www.nch.com.au/scribe/index.html) and [a tutorial](https://www.youtube.com/watch?v=kbKRJcd1ueA). There is a free version.

1. [Transcribe](https://transcribe.wreally.com/)

1. [Nvivo](https://www.qsrinternational.com/nvivo-qualitative-data-analysis-software/about/nvivo/modules/transcription)

1. [Listen N Write](https://download.cnet.com/Listen-N-Write/3000-2170_4-75416316.html)

1. [InScribe](https://www.inqscribe.com/)

1. [Pearnote for Mac](http://www.usefulfruit.com/pearnote/): make text annotations that connect to locations in audio and video recordings (30 day free trial)

1. [Dragon Speech Recognition Software](https://www.nuance.com/dragon.html)

1. [Sonix](https://sonix.ai/)

1. Adobe Premiere: [Speech-to-Text Tutorial](https://digitalanarchy.com/blog/video-editing-tutorial/speech-to-text-using-speech-analysis-in-premiere-pro-creative-cloud/) and [Workaround if you are using a version later than 7.2.2](https://helpx.adobe.com/premiere-pro/using/speech-analysis.html). UF Students can use Adobe products on [UFApps](https://info.apps.ufl.edu/adobeatufapps/)

## Resources & Tools

### Transcription Style & Coding Guides

1. Samuel Proctor Project [Style Guide](https://ufdc.ufl.edu/IR00002513/00001)

1. Library of Congress Guide: [How to Transcribe](https://crowd.loc.gov/help-center/how-to-transcribe/)

1. Smithsonian Archives of American Art [Style Guide](https://www.aaa.si.edu/documentation/oral-history-program-style-guide)
    
1. Smithsonian Volunteer Transcription Program's [Guide](https://transcription.si.edu/instructions) and [Expert Tips](https://transcription.si.edu/articles/insights-tips-volunpeer-expert-transcribe-style)

1. Guide to [Transcription Coding](https://guides.uflib.ufl.edu/celebratingcuba/translation-transcription#s-lg-box-25015800)

1. Hepburn, A., & Bolden, G. B. (2017). Transcribing for social research. Sage. [linkto book descrption on Sage website](https://us.sagepub.com/en-us/nam/transcribing-for-social-research/book237847)

1. Manchester University's very useful [Transcription Toolkit and Guide](http://hummedia.manchester.ac.uk/schools/soss/morgancentre/toolkits/08-toolkit-transcribing-your-qual-data.pdf)

1. Manchester University's [Quick calculator for estimating transcribing time (.xls file)](http://hummedia.manchester.ac.uk/schools/soss/morgancentre/toolkits/08-transcription-planning-calculator.xls)

1. Programming Historian Guide to [Text Automation](https://digital-historian.com/text-automation/)

### Examples - Transcription Notes
    
1. [Sample 1](https://ufdc.ufl.edu/AA00062074/00001/pdf)
    
1. [Sample 2](http://www.gutenberg.org/files/21829/21829-h/21829-h.htm)


### Useful Websites

1. [Linguee](https://www.linguee.com/) Good for slang and expressions;searches different ways in which one word has been translated. Can also enter several words or even a sentence to find different translations.

1. [Thesaurus.com](Thesaurus.com); equivalent in foreign language.

1. Side-by-Side [Text Comparison Website](https://www.diffchecker.com/diff)

1. Jennifer Marie's YouTube channel on [Transcription Tools & Tips](https://www.youtube.com/channel/UC-ooPg34j8PK3osA4l-726Q) 

### Pedals

1. [Infinity](https://www.amazon.com/Infinity-Digital-Control-Computer-USB2/dp/B002MY6I7G/)
    
2. [Kinesis](https://kinesis-ergo.com/products/#foot-switches)

### Text Shortcuts
    
1. [Google Docs](https://www.howtogeek.com/448511/how-to-use-text-shortcuts-in-google-docs/)

1. [MS Word](https://wordribbon.tips.net/T006829_Inserting_Text_with_a_Shortcut_Key.html#:~:text=Position%20the%20insertion%20point%20in,commands%20currently%20using%20this%20shortcut.)

### Misc. Software & Apps

1. [SayMore](https://software.sil.org/saymore/): tools for common Language Documentation tasks

1. [Lameta](https://sites.google.com/site/metadatatooldiscussion/home) tool for metadata associated with files made in the course of documenting language, music, and other cultural expressions.

1. [voice recorder app](https://apps.apple.com/us/app/easy-voice-recorder/id1222784166) for iPad

1. [App for recording phone interviews](https://www.tapeacall.com) (note: be aware of relevant local and national laws regarding recording phone calls).  
  
### Post-Transcription Text Analytics

1. [Planning for text analytics (extraction and organization) in R](https://towardsdatascience.com/organizing-your-first-text-analytics-project-ce350dea3a4a)  

1. Silge and Robinson's **outstanding** book ['Text Analysis in R'](https://www.tidytextmining.com/)

### Anonymization

1. UK Data Service Webinar: [How to anonymise qualitative and quantitative data](https://www.youtube.com/watch?v=Yysj1NaPhls)

1. Anonymizing in R: [part 1](https://www.youtube.com/watch?v=xA2vaUdvxNY) and [part 2](https://www.youtube.com/watch?v=xGvwBMOW4t4)

# Sources / Additional Readings

1. Young, JC, Rose, DC, Mumby, HS, et al. 2018. A methodological guide to using and reporting on interviews in conservation science research. _Methods Ecol Evol_ 9:10-19 [link](https://doi.org/10.1111/2041-210X.12828)

1. Michael Henry Heim & Andrzej W. Tymowski. [Guidelines for the Translation of Social Science Texts](https://www.acls.org/uploadedFiles/Publications/Programs/sstp_guidelines.pdf). _American Council of Learned Societies_

3. Sophie Chabeda, Jane Kahindi, Manya Van Ryneveld. [Preparing data: the not-so-simple stage of transcription and translation](https://globalhealthsocialscience.tghn.org/articles/preparing-data-not-so-simple-stage-transcription-and-translation/)

4. Squires, A. 2009. [Methodological challenges in cross-language qualitative research: A research review.](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2784094/pdf/nihms98173.pdf). *International Journal of Nursing Studies* 46(2):277-287.

5. Hepburn, A., & Bolden, G. B. (2017). [Transcribing for social research](https://uk.sagepub.com/en-gb/eur/transcribing-for-social-research/book237847). Sage.

6. McLellan E, MacQueen KM, Neidig JL. 2003. Beyond the Qualitative Interview: Data Preparation and Transcription. _Field Methods_. 15(1):63-84. [link](https://journals.sagepub.com/doi/10.1177/1525822X02239573)

7. MacLean LM, Meyer M, Estable A. 2004. Improving Accuracy of Transcripts in Qualitative Research. _Qualitative Health Research_.14(1):113-123. [link](https://doi.org/10.1177/1049732303259804)

8. Hughes, M. [Are Online Transcription Services Safe and Private?](https://www.howtogeek.com/657944/are-online-transcription-services-safe-and-private/). _How-to-Geek_.

9. Duca, Daniela. [Who’s disrupting transcription in academia?](https://ocean.sagepub.com/blog/whos-disrupting-transcription-in-academia). _Sage Ocean_.





<!-- *  [Video Interviews](http://brunalab.org/las6291/2018/01/19/video-production/)   -->

<!-- *  snolen says otter is the best: https://otter.ai/login thay have free ed account from wiorecutter: "Otter offers 600 minutes of free transcription per month (as long as you record directly from the app or Zoom), which makes it an attractive option for anyone who wants to transcribe and sort lots of interviews on a nonexistent budget."   -->

<!-- For other types of Recorded Audio:  -->
<!-- Editing and analyzing audio files: https://www.ocenaudio.com/whatis  -->
<!-- *  https://smallbusiness.chron.com/install-speech-office-excel-40728.html -->
<!-- *  https://www.naturalreaders.com/ -->
<!-- # Other Voice to Text -->

<!-- * https://blog.hubspot.com/marketing/voice-to-text  -->

<!-- Guide to capturing and processing [images for archiving](https://guides.uflib.ufl.edu/celebratingcuba/digitization#s-lg-box-wrapper-29381548)   -->

