---
title: "LAS 6292: Reproducibility"
date: "updated: `r Sys.Date()`"
header-includes:
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
- \usepackage[default]{sourcesanspro}
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
# month & year on header left
- \fancyhead[L]{LAS 6292 Spring `r format(Sys.time(), '%Y')`, EM Bruna} 
# name and page on header right
- \fancyhead[R]{Reproducibility, p. \thepage} 
- \fancyfoot[R]{Last update `r format(Sys.time(), '%d %B %Y')`}
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
- \usepackage{xcolor} # changes colors of text
# to center to top level (#) headings (my name) and change color of font
# more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# to center to sub-sub-headings (###) and change color of font
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  # html_document:
  #   df_print: paged
  # rmdformats::robobook: default
---
```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```
# Intro to Reproducibile Data Cleanup in R

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * reminder we will be live-coding

**Bring to Class:  **  

  * supplies to do PB&J Demo
  * post-its in 2 colors (indicating students doing ok or need help)
  * Tent cards for student names
  * info for R users group, uflibrary R 
  * RStudio shortcuts PC or mac
  * Cheat Sheets Rstudio base r,tidyverse, etc.

## Objectives and Competencies:   

  * Understand and be able to explain the importance of a well-documented, reproducible workflow
  * Be able to outline a logical workflow for data importing and correction
  * Be able to load a file as a dataframe in RStudio, edit the data frame, and save the corrected version as a csv file
  

## Pre-class Preparation (Students):

### **Readings**

Read 1-3 before class. #4 goes into more more detail than is necessary before class, but I include it because it will be more useful as you become more experienced.

1. Laskowski, 2020. What to do when you don't trust your data anymore.  [[read online]](https://laskowskilab.faculty.ucdavis.edu/2020/01/29/retractions/) 

<!-- {{% staticref "uploads/readings/Laskowski_2020.pdf" "newtab" %}}[download pdf]{{% /staticref %}} -->

2. Pennisi, E. 2020. Spider biologist denies suspicions of widespread data fraud in his animal personality research. Science. [[read online]](https://www.sciencemag.org/news/2020/01/spider-biologist-denies-suspicions-widespread-data-fraud-his-animal-personality) 
<!-- or {{% staticref "uploads/readings/Pennisi_2020.pdf" "newtab" %}}[download pdf]{{% /staticref %}} -->

3. Alston, J. M., and Rick, J. A.. 2020. A Beginner's Guide to Conducting Reproducible Research. Bull Ecol Soc Am 00( 00):e01801.   [[read online]](https://doi.org/10.1002/bes2.1801) 
<!-- or {{% staticref "uploads/readings/Alston_Rick_2021_BullEcolSocAm.pdf" "newtab" %}}[download pdf]{{% /staticref %}} -->

    **_The following goes into more more detail and will be more useful after you have more practice with R:_**

4.  Wilson G, Bryan J, Cranston K, Kitzes J, Nederbragt L, Teal TK (2017) Good enough practices in scientific computing. PLoS Comput Biol 13(6): e1005510.  [[read online]](https://doi.org/10.1371/journal.pcbi.1005510) 
<!-- or {{% staticref "uploads/readings/Wilson_etal_2017.pdf" "newtab" %}}[download pdf]{{% /staticref %}} -->


### **Online Lectures:**  


1. [A Data Sharing and Management Snafu in 3 Short Acts, by the NYU Health Sciences Library](https://www.youtube.com/watch?v=66oNv_DJuPc)

2. These introductions to R and R Studio were made by Professor Ethan White (UF-WEC). They are a good overview of some R basics.  

      * [Intro to R and RStudio](https://www.youtube.com/watch?v=zqUQL8OOtMQ).
      * [Navigate R and RStudio web page](https://swcarpentry.github.io/r-novice-gapminder/01-rstudio-intro/index.html)
      * [Intro to R Packages](https://www.youtube.com/watch?v=NAArCsmztOI)
      * [Expressions and Variables in R](https://www.youtube.com/watch?v=BFVX0CKY67g&feature=emb_title)
  
**Reminder:** Don't forget to install the `Tidyverse` library; you can read how on [this page](https://tidyverse.tidyverse.org/). And if you want to get ahead to what we will be doing in class, you can read about how to [Set up a Project in R](https://swcarpentry.github.io/r-novice-gapminder/02-project-intro/index.html)

Computer programming is challenging to learn and teach, especially remotely. The next few sessions we will be using modified version of a style known as:  

>`I do it`   
>`We do it`  
>`You do it`

This means you first observe me doing something in Rstudio by watching a video (or in this case you'll watch a video of Prof. Ethan White). The in class we will do some tasks together at the same time (aka "live-coding"). This is useful, because you can ask questions as you go. Then, you will work on an assignment based on the tasks we did together, which you will submit as your assignment.


## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 20
breakout1 <- 15
breakout1_return <-20
snack_break <- 10
proj_overview <- 0
cleaning_demo <- 60

total_time <- 
  t1_overview + 
  breakout1 + 
  breakout1_return + 
  snack_break + 
  proj_overview + 
  cleaning_demo 

time_remain <- 180 - total_time
```


### **Topic 1: Intro to Reproducibility** (`r t1_overview` min)

* Why should we practice ‘Reproducible research’?: 

  <!-- Introduce reproducibility and why it is important; answer any student questions -->

### **Breakout** (`r breakout1` min)

* What is programming? Write the steps necessary for me to build a peanut-butter and jelly sandwich.

<!-- http://static.zerorobotics.mit.edu/docs/team-activities/ProgrammingPeanutButterAndJelly.pdf -->
<!-- https://www.youtube.com/watch?v=HXl5f2azATU  -->

### **Returning results** (`r breakout1_return` min)

* Running your Code: Sandwich or No Sandwich?
* Think like a recipe: what do you need, what are the groups of actions taken?
* Programming is first about logic and organization - step away from the computer and diagram the objectives and steps. 
* Break things into modular components (= 'chunks').
* note any actions or corrections that are repeated multiple times. 

### **Break** (`r snack_break` min)

### **Coding Corrections: Intro** (`r cleaning_demo` min)

### **Coding Corrections: Now do it on your own** (`r time_remain` min)

