---
title: "LAS 6292: Reproducible Data (Re)Organization"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

  * Understand and be able to explain the importance of a well-documented, reproducible workflow
  * Be able to outline a logical workflow for data importing and correction
  * Be able to load a file as a dataframe in RStudio, edit the data frame, and save the corrected version as a csv file
  
  

## Pre-class Preparation (Students):

### **Readings**
1. Laskowski, 2020. What to do when you don't trust your data anymore.  
        [[read online]](https://laskowskilab.faculty.ucdavis.edu/2020/01/29/retractions/) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/laskowski_2020.pdf)

2. Pennisi, E. 2020. Spider biologist denies suspicions of widespread data fraud in his animal personality research. Science.  
        [[read online]](https://www.sciencemag.org/news/2020/01/spider-biologist-denies-suspicions-widespread-data-fraud-his-animal-personality) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Pennisi_2020.pdf)

3.  Alston, J. M., and Rick, J. A.. 2020. A Beginner's Guide to Conducting Reproducible Research. Bull Ecol Soc Am 00( 00):e01801. 
        [[read online]](https://doi.org/10.1002/bes2.1801) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Alston_Rick_2021_BullEcolSocAm.pdf)

4.  Wilson G, Bryan J, Cranston K, Kitzes J, Nederbragt L, Teal TK (2017) Good enough practices in scientific computing. PLoS Comput Biol 13(6): e1005510.  
        [[read online]](https://doi.org/10.1371/journal.pcbi.1005510)  [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/wilson_etal_2017.pdf)


### **Online Lectures:**  

These introductions to R and R Studio were made by Professor Ethan White (UF-WEC). They are a good overview of some R basics. 
* [Intro to R and RStudio Video](https://www.youtube.com/watch?v=zqUQL8OOtMQ). 
* [Navigate R and RStudio web page](https://swcarpentry.github.io/r-novice-gapminder/01-rstudio-intro/index.html)
* [Intro to R Packages](https://www.youtube.com/watch?v=NAArCsmztOI)
* [Expressions and Variables in R](https://www.youtube.com/watch?v=BFVX0CKY67g&feature=emb_title)

**Reminder:** Don't forget to install the `Tidyverse` library; you can read how on [this page](https://tidyverse.tidyverse.org/). And if you want to get ahead to what we will be doing in class, you can read about how to [Set up a Project in R](https://swcarpentry.github.io/r-novice-gapminder/02-project-intro/index.html)


Computer programming is challenging to learn and teach, especially remotely. The next few sessions we will be using modified version of a style known as:
`(1) I do it` 
`(2) We do it`
`(3) You do it`
This means you first observe me doing something in Rstudio by watching a video (or in this case you'll watch a video of Prof. Ethan White). The in class we will do some tasks together at the same time (aka "live-coding"). This is useful, because you can ask questions as you go. Then, you will work on an assignment based on the tasks we did together, which you will submit as your assignment.

## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



### **Topic 1 Overview: -------** (`r t1_overview` min)

**Make some videos of this so they have as a guide for the next steps.**
Set up R studio project, create 3 new folders: `data_dirty`, `data_clean`, `scripts`
download data into `dirty_data` folder
look at data, id mistakes and missing stuff
make a few changes
add column
change to lower
save as csv
sweep the environment and restart r to make sure it works

### **Topic 2 Overview: Topic**  (`r t2_overview` min)

Now do it on your own, with a new dataset
Here is the dataset, and here are the instructions.
can you reproduce the results? don't forget to sweep the environment and restart r

5. Save dirty data in the correct folder
6. Import dirty data

read colnmaes
change column names
get summary of data types str, glimpse

https://www.dataquest.io/blog/load-clean-data-r-tidyverse/ change all column names at once
change Data Types
change to a column to lower 
change a column to upper
order factor
add a column
add a calculation
Export clean data
Save code to folder


Upload to HW


Live code changes to the dataset 

### Breakout & Return Results 

**Breakout (`r breakout1` min): ** topic of breakout 

**Returning results & Take-home message (`r breakout1_return` min)** 
summary of results

2. **Take-home message:** message.

additional text
      
### anything before the break? (`r cleaning_demo` min)

if so, describe here    
    
### Break (`r snack_break` min)

### Anything after the break? (`r proj_overview` min)

If so describe here


### Breakout 2: In-class Assignment (`r breakout2` min)
Text about the breakout


### Free Time

There are `r time_remain` min remaining that can be used to -----


# Tools & Resources

# Sources

