---
title: "LAS 6292: DATA MANAGEMENT PLANS"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* 
* 
  
## Pre-class Preparation (Students):

### **Readings**


### **Online Lectures:**  

1. Atul Gawande's TED Talk: [The Importance & Value of the CHECK LIST](https://www.youtube.com/watch?v=55Nc8nccPa0). Follow-up with the details in this good [animated summary](https://www.youtube.com/watch?v=bCDmDzIOZ-g).

2. Jess Stratton (LinkedIn Learning) Video Overview: [Use Google Forms to Create Surveys](https://www.linkedin.com/learning/google-sheets-advanced-tips-and-tricks/use-google-forms-to-create-surveys?u=41282748) (7 min.)

3. OPTIONAL: A more advanced, multi-video [Google Forms Essential Training Course](https://www.linkedin.com/learning/google-forms-essential-training-2/collect-data-easily-with-google-forms?u=41282748) has short (1-3 min) videos explaining each step in more detail. It's great, and only 39 min long from start to finish.


## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



### **Topic 1 Overview: -------** (`r t1_overview` min)


### *Topic 2 Overview: Topic*  (`r t2_overview` min)


Intro text

### Breakout & Return Results 

**Breakout (`r breakout1` min): ** topic of breakout 

**Returning results & Take-home message (`r breakout1_return` min)** 
summary of results

2. **Take-home message:** message.

additional text
      
### anything before the break? (`r cleaning_demo` min)

if so, describe here    
    
### Break (`r snack_break` min)

### Anything after the break? (`r proj_overview` min)

If so describe here


### Breakout 2: In-class Assignment (`r breakout2` min)
Text about the breakout

look over a paper form<-fix teh form on paper<-build a google form


### Free Time

There are `r time_remain` min remaining that can be used to -----


# Tools & Resources

1. Forms

    *[make a template for data entry (form) with Excel](https://www.pryor.com/blog/ease-the-pain-of-data-entry-with-an-excel-forms-template/)
    
    * Leila Gharani: [Create QR Codes for FREE | Use Anywhere (Excel, Word & PowerPoint)](https://www.youtube.com/watch?v=GadmYWAVnjE)
    
    [How to Create a QR Code and Use It Effectively](https://www.youtube.com/watch?v=z22UChzno9Q)
    
    * [Mass Generate QR Codes to Prefill Google Forms](https://www.youtube.com/watch?v=RS45HVZrnfg)
    
    * [Avery Waterproof labels](https://www.avery.com/category/material/all/?category=waterproof-labels)

    * [How to design survey forms for quick data entry](http://www.tools4dev.org/resources/how-to-design-surveys-and-forms-for-quick-data-entry/). This is a great post with suggestions on designing survey forms for work in an international / multi-language context.

    * [Best Practice for Form Design](https://medium.com/nextux/form-design-best-practices-9525c321d759). These suggestions come from a UX-perspective for online forms, so they might seem most appropriate for online surveys. But the suggetions are really good for paper survey forms as well.
https://www.pinterest.es/nunile/ux-bad-practices/

2. Lists 

    * Paula Rizzo's LinkedIn Course [The Power of Lists](https://www.linkedin.com/learning-login/share?forceAccount=false&redirect=https%3A%2F%2Fwww.linkedin.com%2Flearning%2Fthe-power-of-lists-to-get-stuff-done%3Ftrk%3Dshare_ent_url%26shareId%3DH2pUj81TTkicWyBeXQHdxw%253D%253D&account=41282748)


    * Atul Gawande's [The Checklist Manifesto](http://atulgawande.com/book/the-checklist-manifesto/) 
    
    
3. Apps / Software for Checklists
    
    * Clear
    * Todoist
    * Evernoite
  
  
  Forms
  tapforms (can print labels)
  jotforms
  
  

# Sources

