---
title: "LAS 6292: AUTOMATED DATA EXTRACTION"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

https://conbio.onlinelibrary.wiley.com/doi/10.1111/cobi.14257

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* OCR 

## Pre-class Preparation (Students):

### **Readings**

### **Online Lectures:**  

## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```

### **Topic 1 Overview: Optical Character Recognition (OCR)** (`r t1_overview` min)
1. 3 folders with 9 photos, save to google drive

OCR to text - try doing them all, handwriting is obviously the messiest one, try doing it with your own. lets all write a text, and then try to see whow it does.  

no way to go directly to google sheets, need tihighlight (leave blue) to to threre icos, then open with google docs. download as text

https://www.pdf2go.com/
compare with: https://smallpdf.com/ 


1. What is OCR?

1. OCR: google, website, R
2. Data from Published Figs
3. Data Packages Govt Data via API
4. text analysis


### *Topic 2 Overview: Topic*  (`r t2_overview` min)

Image is from: Pereira, Thales Augusto Zamberlan. (2018). Poor Man's Crop? Slavery in Brazilian Cotton Regions (1800-1850). Estudos Econômicos (São Paulo), 48(4), 623-655. https://doi.org/10.1590/0101-41614843tzp

Intro text

### Breakout & Return Results 

(we did this in-class together as a live-coding-type-exercise): using * [Web Plot Digitizer](https://automeris.io/WebPlotDigitizer/), we extracted data from several figures that differ in quality and content.


**Breakout (`r breakout1` min): ** topic of breakout 

**Returning results & Take-home message (`r breakout1_return` min)** 
summary of results

2. **Take-home message:** message.

additional text
      
### anything before the break? (`r cleaning_demo` min)

if so, describe here    
    
### Break (`r snack_break` min)

### Free Time

There are `r time_remain` min remaining that can be used to -----
web

