---
title: "LAS 6292: DATA VISUALIZATION"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* 
  
## Pre-class Preparation (Students):

### **Readings**

1. [“Ch7 in Exploratory Data Analysis” in Wickham and Grolemund’s “R for Data Science”:](https://r4ds.had.co.nz/exploratory-data-analysis.html)  

2. Rougier NP, Droettboom M, Bourne PE (2014) Ten Simple Rules for Better Figures. PLoS Comput Biol 10(9): e1003833. [read online](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003833)



Original Readings for this week: 

1.  Alexander, S.M., Jones, K., Bennett, N.J. et al. Qualitative data sharing and synthesis for sustainability science. Nat Sustain 3, 81–88 (2020).  
        [[read online]](https://doi.org/10.1038/s41893-019-0434-8 )  [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Alexander_etal_2019_NatureSustainability.pdf)

2. Renaut, S. et al. 2018. Management, Archiving, and Sharing for Biologists and the Role of Research Institutions in the Technology-Oriented Age. *BioScience* 68(6)400–411  
        [[read online]](https://academic.oup.com/bioscience/article/68/6/400/4983937) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Renaut_etal_2018_Bioscience.pdf)  

2.  Duke, Clifford S., and John H. Porter. 2013. "The ethics of data sharing and reuse in biology." BioScience 63(6): 483-489.  
        [[read online]](https://academic.oup.com/bioscience/article/63/6/483/226339) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Duke%26Porter_2013.pdf)

3.  Tenopir, C., et al.  (2015). Changes in data sharing and data reuse practices and perceptions among scientists worldwide. PloS one, 10(8), e0134826.  
        [[read online]](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0134826) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Tenopir_etal_2015.PDF)

4.  Mauthner, NS, & O. Parry O. 2009. Qualitative data preservation and sharing in the social sciences: On whose philosophical terms?. Australian J of Social Issues 44(3):291-307.  
         [[read online]](https://onlinelibrary.wiley.com/doi/abs/10.1002/j.1839-4655.2009.tb00147.x)  [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Mauthner_et_al-2009-Australian_Journal_of_Social_Issues.pdf)


### **Online Lectures:**  

1. Tommy McCall: [The simple genius of a good graphic](https://www.ted.com/talks/tommy_mccall_the_simple_genius_of_a_good_graphic?language=en)
    
2. R. Luke DuBois: [Insightful human portraits made from data](https://www.ted.com/talks/r_luke_dubois_insightful_human_portraits_made_from_data?language=en)

3. Mike Chapple: [Data Visualization in R with ggplot2](https://www.linkedin.com/learning/data-visualization-in-r-with-ggplot2/the-grammar-of-graphics?resume=false&u=41282748). Watch the "Introducing ggplot2" (4:43) and "Grammar of Graphics" (4:54)


  
## Class Outline
https://datacarpentry.org/R-ecology-lesson/04-visualization-ggplot2.html
for lecture https://www.yan-holtz.com/PDF/Dataviz_Caveats.pdf


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



