---
title: "LAS 6292: Data, Ethics, and the Law"
date: "updated: `r Sys.Date()`"
header-includes:
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
- \usepackage[default]{sourcesanspro}
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
# month & year on header left
- \fancyhead[L]{LAS 6292 Spring `r format(Sys.time(), '%Y')`, EM Bruna} 
# name and page on header right
- \fancyhead[R]{Data Ethics, and the Law, p. \thepage} 
- \fancyfoot[R]{Last update `r format(Sys.time(), '%d %B %Y')`}
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
- \usepackage{xcolor} # changes colors of text
# to center to top level (#) headings (my name) and change color of font
# more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# to center to sub-sub-headings (###) and change color of font
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# - \usepackage{ragged2e}
# - \setlength\RaggedRightParindent{15pt}
# - \RaggedRight - # make the right margin ragged
# https://tex.stackexchange.com/questions/258714/prevent-hyphenation-across-lines-enforce-right-margin
# - \hyphenpenalty=10000 #to suppress ordinary hyphenation
# - \exhyphenpenalty=10000 #to prohibit line breaks at explicit hyphens.
# - \setstretch{1.0}

# https://babichmorrowc.github.io/post/changing-fonts-in-rmarkdown-pdfs/

mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  # pdf_document: default
  html_document: defauilt
  #   df_print: paged
  # rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
knitr::opts_chunk$set(echo = TRUE)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## In-Class Exercise: Data, Ethics, & the Law

### PART 1: Anonymizing is hard

1. Exercise 1: Finalize this "anonymization log' for one of your interviews. How would you anonymiuze each entry? 


```{r anon_log, echo=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)

col1<-c(
"Int1 p1",
"Int1 p1",
"Int1 p2",
"Int1 p2",
"Int1 p3",
"Int1 p4",
"Int2 p1",
"Int2 p8",
"Int2 p10",
"Int2 p11"
)

col2<-c(
"Age: 27",
"Born in Venezeula",
"Born on 20 June",
"Name: Miguel",
"Lives in Atlanta",
"Known as 'Tio Miguelito'",
"his friend Maria",
"attended Talbot Elementary School",
"Director, Southern Region, Quantum Vehicles",
"lives in Grant Park"
)

col3<-c(
" ",
" ",
" ",
" ",
" ",
" ",
" ",
" ",
" ",
" "
)

log<-data_frame(col1,col2,col3)
names(log)<-c("Interview-Page","Original Info", "Changed to")

log %>%
  kbl() %>%
  kable_styling(latex_options = "HOLD_position")
```


2. Exercise 2: [Download this interview file](https://github.com/BrunaLab/LAS6292_DataManagement/blob/70a15b1b37357f7dabc4a3c77787cca308d2ba18/content/course-materials/class-sessions/14-legal-ethical/in-class_resources/corti_03_exercise_8.1_original_answer.docx) by clicking the 'download" button and saving to your computer. Try anonymizing this text - was it harder or easier than the more abstract list above? Why or why not?


### PART 2: Research misconduct related to data

1. Take 15 minutes to read this article: Martinson, B., Anderson, M. & de Vries, R. 2005. Scientists behaving badly. Nature 435:737–738. [link](https://doi.org/10.1038/435737a). We'll come back and discuss it as a group aftyer the break.

