---
title: "LAS 6292: TOPIC"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

  *Be able to explain what metadata are and why they are important
  *Identify & list the types of information typically included in metadata records
  *Identify reasons metadata is of value to data users, creators, and organizations
  *Be able to identify metadata standards relevant to different disciplines
  * Explain the elements of good metadata


## Pre-class Preparation (Students):

### **Readings**

1.  Michener, W.K., et al . 1997. Non-geospatial metadata for the ecological sciences. Ecological Applications 7: 330–342.  
        [[read online]](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1890/1051-0761%281997%29007%5B0330%3ANMFTES%5D2.0.CO%3B2) [[download pdf]](https://github.com/BrunaLab/LAS6292_DataManagement/blob/master/Assigned_Readings/Michener_etal_1997_EcolApplications.pdf)

2.  Pp 446-450 in Bernard, H.R. and Bernard, H.R., 2013. Social research methods: Qualitative and quantitative approaches. Sage.  
        [[download pdf]]


### **Online Lectures:**  

* What are Metadata

  
## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```



### **Topic 1 Overview: -------** (`r t1_overview` min)

Any questions?

### **Topic 2 Overview: Topic**  (`r t2_overview` min)


### Breakout & Return Results 

**Breakout (`r breakout1` min): ** topic of breakout 

Review the Metadata of Others. Did they include the 5 metadata descriptors?

**Returning results & Take-home message (`r breakout1_return` min)** 

2. **Take-home message:** message.

Metadata are hard to write but invaluable
      
### anything before the break? (`r cleaning_demo` min)

<!-- if so, describe here     -->
### Breakout 2: After the Break (`r snack_break` min), In-class Assignment (`r breakout2` min)

Work on your metadata: you can either find


### Free Time

There are `r time_remain` min remaining that can be used to -----

# Tools & Resources

## Codebooks

1. ICPSR: [What is a Codebook?](https://www.icpsr.umich.edu/icpsrweb/content/shared/ICPSR/faqs/what-is-a-codebook.html)

## Organizations Developing Metadata Standards and Schema 

1. [The Research Data Alliance (RDA) ](https://www.rd-alliance.org/)"has the goal goal of building the social and technical infrastructure to enable open sharing and re-use of data."

2. [DDI Alliance](https://ddialliance.org/about/about-the-alliance): "Established in 2003, the Data Documentation Initiative Alliance (DDI Alliance) is an international collaboration dedicated to establishing metadata standards and semantic products for describing social science data, data covering human activity, and other data based on observational methods." 

3. T[he Dublin Core Metadata Initiative](https://www.dublincore.org/) is "an organization supporting innovation in metadata design and best practices across the metadata ecology". 

4. ["Best Practices in Creating Social Science Metadata."](https://www.icpsr.umich.edu/files/deposit/dataprep.pdf) p.32 in the ICPSR *Guide to Social Science Data Preparation and Archiving: 
Best Practice Throughout the Data Life Cycle (6th Edition)*.

## Discipline-Specific Metadata Standards

1. [RDA Catalog of metadata standards for different disciplines](https://rd-alliance.github.io/metadata-directory/subjects/)

2. Ecological data: [Ecological Metadata Language](http://knb.ecoinformatics.org/eml_metadata_guide.html)

3. Museum Specimens: [Darwin Core](http://rs.tdwg.org/dwc/index.htm)

4. Geography Markup Language ([GML](http://www.opengeospatial.org/standards/gml)): Emphasis on geographic features (roads, highways, bridges)


## Tools for Creating machine-readable metadata 

1. [giant list from the RDA of tools for creating standardized metadata for different disciplines ](https://rd-alliance.github.io/metadata-directory/tools/)

2. [USGS Metadata Wizard](https://www.sciencebase.gov/catalog/item/50ed7aa4e4b0438b00db080a)

3. [TKME](https://geology.usgs.gov/tools/metadata/tools/doc/tkme.html)

4. [CatMDEdit](http://catmdedit.sourceforge.net/)

5. [GRIIDC](https://data.gulfresearchinitiative.org/metadata-editor-start)

## Metadata Dictionaries

1. [USGS](https://apps.usgs.gov/thesaurus/)

2. [Global Change Master Dictionary](https://idn.ceos.org/)

3. [USGS Geographic Names](https://www.usgs.gov/core-science-systems/ngp/board-on-geographic-names)

4. [Getty Thesaurus of Gheographic Names](http://www.getty.edu/research/tools/vocabularies/tgn/)

# Sources

1. ICPSR [*Guide to Social Science Data Preparation and Archiving: 
Best Practice Throughout the Data Life Cycle (6th Edition)*](https://www.icpsr.umich.edu/web/pages/deposit/guide/)

2. DataONE Community Engagement & Outreach Working Group (2017) "Metadata Management". Accessed through the Data Management Skillbuilding Hub at https://dataoneorg.github.io/Education/lessons/07_metadata/index on Aug 31, 2020

