---
title: "LAS 6292: AUTOMATED DATA EXTRACTION"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: blue
urlcolor: blue
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Pre-Class Preparation (Instructor):  

**Send in an email to students:  **  

  * content of any pre-class emails.

**Bring to Class:  **  

  * Snacks
  * Flip charts and markers
  * Dry write markers
  * Tent cards for student names

## Objectives and Competencies:   

* OCR 

## Pre-class Preparation (Students):

### **Readings**

### **Online Lectures:**  

## Class Outline


```{r timer, echo=FALSE}
t1_overview <- 10
t2_overview <- 10
breakout1 <- 15
breakout1_return <-35
cleaning_demo <-10
snack_break <- 10
proj_overview <- 15
breakout2 <- 45

total_time <- t1_overview + 
  t2_overview + 
  breakout1 + 
  breakout1_return + 
  cleaning_demo +
  snack_break + 
  proj_overview + 
  breakout2 

time_remain <- 180 - total_time
```

### **Topic 1 Overview: Optical Character Recognition (OCR)** (`r t1_overview` min)
1. 3 folders with 9 photos, save to google drive

OCR to text - try doing them all, handwriting is obviously the messiest one, try doing it with your own. lets all write a text, and then try to see whow it does.  

no way to go directly to google sheets, need tihighlight (leave blue) to to threre icos, then open with google docs. download as text

https://www.pdf2go.com/
compare with: https://smallpdf.com/ 


1. What is OCR?

1. OCR: google, website, R
2. Data from Published Figs
3. Data Packages Govt Data via API
4. text analysis


### *Topic 2 Overview: Topic*  (`r t2_overview` min)

Image is from: Pereira, Thales Augusto Zamberlan. (2018). Poor Man's Crop? Slavery in Brazilian Cotton Regions (1800-1850). Estudos Econômicos (São Paulo), 48(4), 623-655. https://doi.org/10.1590/0101-41614843tzp

Intro text

### Breakout & Return Results 

(we did this in-class together as a live-coding-type-exercise): using * [Web Plot Digitizer](https://automeris.io/WebPlotDigitizer/), we extracted data from several figures that differ in quality and content.


**Breakout (`r breakout1` min): ** topic of breakout 

**Returning results & Take-home message (`r breakout1_return` min)** 
summary of results

2. **Take-home message:** message.

additional text
      
### anything before the break? (`r cleaning_demo` min)

if so, describe here    
    
### Break (`r snack_break` min)

### Free Time

There are `r time_remain` min remaining that can be used to -----


# Tools & Resources


# Automated Data Collection & Extraction

## Tools & Resources

### Optical Character Recognition

1.Video Primer: [What is OCR?](what is ocr? https://www.youtube.com/watch?v=Q5U_VEmcY-M)

2. Online OCR Tools - text and data from pdfs into csv,txt, etc.

  * Google Drive - Video Primer: [OCR with Google Drive](https://www.youtube.com/watch?v=4lApeCNK6Ic)
  * Free online sites for small batches (can upgrade for larger numbers of files)
    
      * [Free Online OCR 1](http://www.free-online-ocr.com/)
      * [New OCR](https://www.newocr.com/)
      * [pdf to excel](https://smallpdf.com/pdf-to-excel)
      * [OnlineOCR](https://www.onlineocr.net/)
      * [PDFTables](https://pdftables.com/) will convert PDF to .csv, and has an [API](https://pdftables.com/blog/convert-pdf-to-excel-r) so you can do your conversions in bulk with R. You can do ~25 pages free; large numbers are reasonably priced.
  
  * [Mathpix Snip](https://mathpix.com/) digitizes handwritten or printed text, and copies outputs to the clipboard that can be pasted into LaTeX editors like Overleaf, Markdown editors like Typora, Microsoft Word, and more. 
  
3. R tools for OCR: `pdfreader`, `tabulizer`, and `pdfreader`
 * R package [`pdftools`](https://cran.r-project.org/web/packages/pdftools/index.html)
 * [written tutorial 1](http://theautomatic.net/2018/08/24/getting-data-from-pdfs-the-easy-way-with-r/)
 * [written tutorial 2](https://www.charlesbordet.com/en/extract-pdf/#)
 * [written tutorial 3](https://datascienceplus.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)
 * [Video Tutorial 1](https://www.youtube.com/watch?v=DhY3V4LCdps)
 * [Video Tutorial 2](https://www.youtube.com/watch?v=6ItyOgS2noc)
 * Detailed [Blog Post / Tutorial](https://datascienceplus.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)
  * [Convert PDF to text in R OCR pdftools](https://www.youtube.com/watch?v=DhY3V4LCdps)
  * [PDFtools in R](https://www.youtube.com/watch?v=6ItyOgS2noc )
    

4. Extracting Tables from images using R package `magick`.
  
  * Detailed [Blog Post / Tutorial](https://themockup.blog/posts/2021-01-18-reading-tables-from-images-with-magick/)

### Extracting Data from Published Figures

1. Ankit Rohagni's [Web Plot Digitizer](https://automeris.io/WebPlotDigitizer/)
   * WPD [Video Tutorial](https://www.youtube.com/watch?v=P7GbGdMvopU)
   * WPD Tutorial [Blog Post](https://towardsdatascience.com/extracting-digitising-data-from-plots-in-scientific-papers-or-images-6e098cb887aa)  

1. Alternative 1: R package [`magick`](https://www.r-bloggers.com/2019/06/extracting-the-data-from-static-images-of-graphs-with-magick/)  

1. Alternative 2: [GetData](http://www.getdata-graph-digitizer.com/index.php) extracts data automatically from scanned images (~$30).  

1. Alternative 3: R package [`digitize`](https://cran.r-project.org/web/packages/digitize/digitize.pdf) will extract data from scatterplots within the R environment. [This article](https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Poisot.pdf) will walk you through the process.

### Text Mining

1. [Text Mining with R](https://www.tidytextmining.com/) by Julia Silge and David Robinson

1. [`gutenbergr`](https://cran.r-project.org/web/packages/gutenbergr/index.html): Download and Process Public Domain Works from [Project Gutenberg](https://www.gutenberg.org/). Tutorial can be found [here](https://docs.ropensci.org/gutenbergr/)

1. Atanassova I, Bertin M and Mayr P (2019) Editorial: Mining Scientific Papers: NLP-enhanced Bibliometrics. Front. Res. Metr. Anal. 4:2. [doi: 10.3389/frma.2019.00002](https://www.frontiersin.org/articles/10.3389/frma.2019.00002/full)

1. Westergaard D, Stærfeldt H-H, Tønsberg C, Jensen LJ, Brunak S (2018) A comprehensive and quantitative comparison of text-mining in 15 million full-text articles versus their corresponding abstracts. PLoS Comput Biol 14(2): e1005962. https://doi.org/10.1371/journal.pcbi.1005962

1. Salloum, S.A., Al-Emran, M., Monem, A.A., Shaalan, K. (2018). Using Text Mining Techniques for Extracting Information from Research Articles. In: Shaalan, K., Hassanien, A., Tolba, F. (eds) Intelligent Natural Language Processing: Trends and Applications. Studies in Computational Intelligence, vol 740. Springer, Cham. https://doi.org/10.1007/978-3-319-67056-0_18

1. Simon, C., Davidsen, K., Hansen, C. et al. BioReader: a text mining tool for performing classification of biomedical literature. BMC Bioinformatics 19, 57 (2019). https://doi.org/10.1186/s12859-019-2607-x
Extracting Body Text from Academic PDF Documents for Text Mining

1. Benchimol, J., Kazinnik, S., & Saadon, Y. (2022). Text mining methodologies with R: An application to central bank texts. Machine Learning with Applications, 8, 100286.[link](https://scholar.harvard.edu/files/jbenchimol/files/text-mining-methodologies.pdf) 

1. Yu, C., Zhang, C., & Wang, J. (2020). Extracting Body Text from Academic PDF Documents for Text Mining. arXiv preprint arXiv:2010.12647.

1. Gulo, C. A., & Rúbio, T. R. (2015, January). Text Mining Scientific Articles using the R. In Doctoral Symposium in Informatics Engineering. [linl](https://paginas.fe.up.pt/~prodei/dsie15/web/papers/dsie15_submission_10.pdf)

1. [jstor: An R package for Analysing Scientific Articles](https://www.youtube.com/watch?v=kNRbT-ki9tU)

1. Processing Records from the WOS and Scopus for Text Mining & Biliometrics

* R package [`refsplitr`](https://github.com/ropensci/refsplitr)

* R package [`bibliometrix`](https://www.bibliometrix.org/)

### Web Scraping

1. Scraping information from websites 
  * Noortje Marres & Esther Weltevrede (2013) Scraping the Social?, Journal of Cultural Economy, 6:3, 313-335, DOI: 10.1080/17530350.2013.772070

  * Library Carpentry [Lesson on Webscraping](https://librarycarpentry.org/lc-webscraping/) 
  * Start Here: [Introduction to webscraping](https://nceas.github.io/oss-lessons/data-liberation/intro-webscraping.html)
  * Video: [Scraping WebData in R with rvest ](https://www.youtube.com/watch?v=4IYfYx4yoAI)
  * Video: [Practical Introduction to Web Scraping using R](https://www.youtube.com/watch?v=l37n_HDD1qs)
  * Very nice [written tutorial](https://www.scrapingbee.com/blog/web-scraping-r/)...
  * ....and another one, this time from the [UC Business Analytics R Programming Guide](http://uc-r.github.io/scraping)
  * [scraping HTML text](http://bradleyboehmke.github.io//2015/12/scraping-html-text.html#specific_nodes) and [scraping HTML tables](http://bradleyboehmke.github.io/2015/12/scraping-html-tables.html)
  * [SelectorGadget](https://selectorgadget.com/) is useful to id CSS selectors

2. Social Media Data

1. Scraping [Twitter Data with R](https://utstat.toronto.edu/~nathan/teaching/sta4002/Class1/scrapingtwitterinR-NT.html) or with [Tweetsets](https://programminghistorian.org/en/lessons/beginners-guide-to-twitter-data)

### Cell Phone Data

1. Exploratory analyses [Part 1](https://www.r-bloggers.com/2019/11/exploratory-data-analysis-of-cell-phone-usage-with-r-part-1/) and [Part 2](https://www.r-bloggers.com/2019/12/exploratory-data-analysis-of-cell-phone-usage-with-r-part-2/)

### Automated Image Analysis

1. Pennekamp, F. and Schtickzelle, N. (2013), Implementing image analysis in laboratory‐based experimental systems for ecology and evolution: a hands‐on guide. Methods Ecol Evol, 4: 483-492. https://doi.org/10.1111/2041-210X.12036 

2. How to build your own image recognition app with R! [Part 1](https://www.r-bloggers.com/2021/03/how-to-build-your-own-image-recognition-app-with-r-part-1/?fbclid=IwAR3P8GNeEHAtm62iOIhhFe45ofZjn-zlwezc5TtM0pbayfUdjmwAr3UoB1Q) and [Part 2](https://www.r-bloggers.com/2021/03/how-to-build-your-own-image-recognition-app-with-r-part-2/)

### Wearable Devices & RFID tags

1. [What is an RFID tag?](https://en.wikipedia.org/wiki/Radio-frequency_identification)

1. Rafiq, K., Appleby, R. G., Edgar, J. P., Radford, C., Smith, B. P., Jordan, N. R., Dexter, C. E., Jones, D. N., Blacker, A. R. F., & Cochrane, M. (2021). WildWID: An open-source active RFID system for wildlife research. Methods in Ecology and Evolution, 12, 1580– 1587. https://doi.org/10.1111/2041-210X.13651

1. [Build your own RFID device](https://news.mongabay.com/2015/07/build-your-own-rfid/)

1. Izmailova, E.S., Wagner, J.A. and Perakslis, E.D. (2018), Wearable Devices in Clinical Trials: Hype and Hypothesis. Clin. Pharmacol. Ther., 104: 42-52. https://doi.org/10.1002/cpt.966

2. Loncar-Turukalo T, Zdravevski E, Machado da Silva J, Chouvarda I, Trajkovik V. Literature on Wearable Technology for Connected Health: Scoping Review of Research Trends, Advances, and Barriers J Med Internet Res 2019;21(9):e14017 [doi: 10.2196/14017](https://www.jmir.org/2019/9/e14017/)

1. [Why Should Sociologists Care about Wearable Tech?](https://www.asanet.org/why-should-sociologists-care-about-wearable-tech)

1. Harari, G. M., Lane, N. D., Wang, R., Crosier, B. S., Campbell, A. T., & Gosling, S. D. (2016). Using Smartphones to Collect Behavioral Data in Psychological Science: Opportunities, Practical Considerations, and Challenges. Perspectives on Psychological Science 11(6), 838-854 [link](https://pubmed.ncbi.nlm.nih.gov/27899727/)

1. Seifert Alexander, Hofer Matthias, Allemand Mathias. 2018. Mobile Data Collection: Smart, but Not (Yet) Smart Enough. 12. Frontiers in Neuroscience https://www.frontiersin.org/article/10.3389/fnins.2018.00971     

### Build your own automated data collection tool, Rasberry PI

*  Calipers that dump data directly to Excel [link](https://www.notion.so/Hacking-Digital-Calipers-3ee7726f11ca431694dc70a1977516e4)  

* [PiSpy](https://www.biorxiv.org/content/10.1101/2022.03.21.485129v1): An Affordable, Accessible, and Flexible Imaging Platform for the Automated Observation of Organismal Biology and Behavior

* Jolles, J. W. (2021). Broad-scale applications of the Raspberry Pi: A review and guide for biologists. Methods in Ecology and Evolution, 12, 1562– 1579. https://doi.org/10.1111/2041-210X.13652

### Databases and Websites from which you can extract data 

1. Government 

  * [Data.gov](https://www.data.gov/) (the open data portal of the US Government) and [Using Data.gov APIs in R](https://data.library.virginia.edu/using-data-gov-apis-in-r/)
  * the [rOpengov Project](http://ropengov.org/)
  * [Open Fiscal Data Package](http://www.fiscaltransparency.net/ofdp/) 
  * [`educationdata`](https://urbaninstitute.github.io/education-data-package-r/#educationdata): Retrieve data from the Urban Institute’s Education Data API as a data.frame for easy analysis. See [also here](https://github.com/UrbanInstitute/education-data-package-r)
  * a [huge list](https://cengel.github.io/gearup2016/SULdataAccess.html) of data sources for social scientists available with R tools 
  * accessing [World bank Data](https://vincentarelbundock.github.io/WDI/) with R

1. US & World Census Data  
  
  * [A Guide to Working with US Census Data in R](https://rconsortium.github.io/censusguide/r-packages-all.html)
  * R Package [`tidycensus`](https://cran.r-project.org/web/packages/tidycensus/tidycensus.pdf)
  * [Tutorial 1](https://walkerke.github.io/tidycensus/articles/basic-usage.html)
  * [Tutorial 2](http://www.computerworld.com/article/3120415/data-analytics/how-to-download-new-census-data-with-r.html rel="nofollow noopener" href="https://hrecht.github.io/censusapi/)
  * R package [`ipumsr`](https://tech.popdata.org/ipumsr/): The ipumsr package helps import IPUMS extracts from the IPUMS website into R. IPUMS provides census and survey data from around the world integrated across time and space. 


1. Education Data 

  * [`edbuildr`](https://github.com/EdBuild/edbuildr): import [EdBuild’s master dataset](http://viz.edbuild.org/workshops/edbuildr/) of school district finance, student demographics, and community economic indicators for every school district in the United States.

  * [Building R and Stata packages for the Education Data Portal](https://urban-institute.medium.com/building-r-and-stata-packages-for-the-education-data-portal-b88760c07233)

1. Other Online Data Sets & Tools for Accessing them

  * Giant [compendium of open datasets #1](https://github.com/awesomedata/awesome-public-datasets)

  * [Data on Amazonia](https://datazoomamazonia.com.br/)
  
<!-- https://github.com/Chicago/RSocrata/">RSocrata GitHub repo</a></td> -->

  * R package [`bdc`](https://brunobrr.github.io/bdc/): toolkit for gathering & cleaning biodiversity data

  * [EcoRetriever](https://retriever.readthedocs.io/en/latest/):  automates the tasks of finding, downloading, and cleaning up publicly available ecological data, and then stores them in a local database or csv files. 

  * [litsearcher](https://elizagrames.github.io/litsearchr/)
an R package to facilitate quasi-automatic search strategy development for systematic review

1. Correia, R.A., Ladle, R., Jarić, I., Malhado, A.C.M., Mittermeier, J.C., Roll, U., Soriano‐Redondo, A., Veríssimo, D., Fink, C., Hausmann, A., Guedes‐Santos, J., Vardi, R. and Di Minin, E. (2021), Digital data sources and methods for conservation culturomics. Conservation Biology, 35: 398-411. https://doi.org/10.1111/cobi.13706


